Characterizing Execution Times on Realistic Programs

### 2018-04-12 pt_characterization/real_data/paper.tex
New draft

The following section shows histograms for runs on 
different real-world programs with varying input sizes.

\section{Histograms of the Execution Times on Real-World Programs~\label{sec:real-world}}

In this section we run three kinds of programs: {\it nested for-loop}, {\it insertion sort} and {\it matrix multiplication}. For the runs of the latter two programs, we varied their input sizes by 2{\small $\times$} while trying some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size. We then exhibit different kinds of histograms in the subsequent sections.

@@

In this section we check if the use of different machines affects the distribution of execution (process) times on the same program. 
In other words, we examine {\em inter-machine repeatability}. For this examination, we use the nested for loop program 
with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.

Figures~\ref{fig:dm_1}, ~\ref{fig:dm_2}, and ~\ref{fig:dm_3} display the process time (PT) histograms on INC16, INC13, and INC17.2 run on our machines (from {\tt sodb8} to {\tt sodb12}), respectively.

@@

\subsection{One-to-One Comparison between An Insertion Sort \& A Corresponding INC Programs~\label{sec:sort}} 
This section compares program time histograms 
of insertion sort and INC (a nested-for-loop) programs. 
The insertion sort program 
sorts the elements of a given array in non-decreasing order.  
The program repeatedly runs 300 times for a given input size. 
The input size for the program varies from 144K to 344K integer elements, which are randomly generated. 
Note that each sort program over a specific input size is termed SORT{\it x}: 
for instance, SORT100 indicates the insertion sort program over 100K elements. 
An INC program's task length is correspondingly determined by the program time 
of an SORT program. In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3} 
we perform a match on an SORT program and its corresponding INC program.

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT144]{
		\includegraphics[scale=0.43]{sort144_dist.eps}
		\label{fig:sort144_dist}
	}	
	\subfigure[PT frequency on INC8.8]{
		\includegraphics[scale=0.43]{inc8.8_dist.eps}
		\label{fig:inc8.8_dist}
	}
	\subfigure[PT frequency on SORT177]{
		\includegraphics[scale=0.43]{sort177_dist.eps}
		\label{fig:sort177_dist}
	}	
	\subfigure[PT frequency on INC13]{
		\includegraphics[scale=0.43]{inc13.0_dist.eps}
		\label{fig:inc13_dist}
	}
	\caption{PT Histograms I~\label{fig:sort1}}
\end{figure}

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT144]{
		\includegraphics[scale=0.43]{sort244_dist.eps}
		\label{fig:sort144_dist}
	}	
	\subfigure[PT frequency on INC25.8]{
		\includegraphics[scale=0.43]{inc25.8_dist.eps}
		\label{fig:inc25.8_dist}
	}
	\subfigure[PT frequency on SORT288]{
		\includegraphics[scale=0.43]{sort177_dist.eps}
		\label{fig:sort177_dist}
	}	
	\subfigure[PT frequency on INC35]{
		\includegraphics[scale=0.43]{inc13.0_dist.eps}
		\label{fig:inc35_dist}
	}
	\caption{PT Histogram Comparison II~\label{fig:sort2}}
\end{figure}

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT344]{
		\includegraphics[scale=0.43]{sort344_dist.eps}
		\label{fig:sort344_dist}
	}	
	\subfigure[PT frequency on INC51.5]{
		\includegraphics[scale=0.43]{inc51.5_dist.eps}
		\label{fig:inc51.5_dist}
	}
	\caption{PT Histogram Comparison III~\label{fig:sort3}}
\end{figure}

@@

\section{Additional Insertion Sort~\label{sec:add_new}}

\begin{figure}[h]
	\centering
%	\subfigure[PT frequency on SORT25]{
%		\includegraphics[scale=0.43]{sort25_dist.eps}
%		\label{fig:sort25_dist}
%	}
%	\subfigure[PT frequency on SORT50]{
%		\includegraphics[scale=0.43]{sort50_dist.eps}
%		\label{fig:sort50_dist}
%	}
	\subfigure[PT frequency on SORT100]{
		\includegraphics[scale=0.43]{sort100_dist.eps}
		\label{fig:sort100_dist}
	}
	\subfigure[PT frequency on SORT400]{
		\includegraphics[scale=0.43]{sort400_dist.eps}
		\label{fig:sort400_dist}
	}
	\subfigure[PT frequency on SORT580]{
		\includegraphics[scale=0.43]{sort580_dist.eps}
		\label{fig:sort588_dist}
	}
	\caption{PT Histograms of SORT100 ... SORT580~\label{fig:n_sort1}}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT800]{
		\includegraphics[scale=0.43]{sort800_dist.eps}
		\label{fig:sort800_dist}
	}
	\subfigure[PT frequency on SORT1000]{
		\includegraphics[scale=0.43]{sort1000_dist.eps}
		\label{fig:sort1000_dist}
	}
	\subfigure[PT frequency on SORT1160]{
		\includegraphics[scale=0.43]{sort1160_dist.eps}
		\label{fig:sort1160_dist}
	}
	\subfigure[PT frequency on SORT2320 (not complete)]{
		\includegraphics[scale=0.43]{sort2320_dist.eps}
		\label{fig:sort2320_dist}
	}
	\caption{PT Histograms of SORT800 ... SORT2320~\label{fig:n_sort2}}
\end{figure}

@@

Figure~\ref{fig:matc} shows a series of 
histograms of the execution times measured on 
the same matrix multiplication program in column major 
as the input sizes grows.

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on MATC1K]{
		\includegraphics[scale=0.43]{matc1024_dist.eps}
		\label{fig:matc1k_dist}
	}
	\subfigure[PT frequency on MATC2K]{
		\includegraphics[scale=0.43]{matc2048_dist.eps}
		\label{fig:matc2k_dist}
	}
	\subfigure[PT frequency on MATC4K]{
		\includegraphics[scale=0.43]{matc4096_dist.eps}
		\label{fig:matc4k_dist}
	}
	\subfigure[PT frequency on MATC8K]{
		\includegraphics[scale=0.43]{matc8192_dist.eps}
		\label{fig:matc8k_dist}
	}
	\caption{PT Histograms of MATC1024 ... MATC8192~\label{fig:matc}}
\end{figure}

@@

Figure~\ref{fig:matc} shows a series of 
histograms of the execution times measured on 
the same matrix multiplication program in row major as the input sizes grows.

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on MATR1K]{
		\includegraphics[scale=0.43]{matr1024_dist.eps}
		\label{fig:matr1k_dist}
	}
	\subfigure[PT frequency on MATR2K]{
		\includegraphics[scale=0.43]{matr2048_dist.eps}
		\label{fig:matr2k_dist}
	}
	\subfigure[PT frequency on MATR4K]{
		\includegraphics[scale=0.43]{matr4096_dist.eps}
		\label{fig:matr4k_dist}
	}
	\subfigure[PT frequency on MATR8K]{
		\includegraphics[scale=0.43]{matr8192_dist.eps}
		\label{fig:matr8k_dist}
	}
	\caption{PT Histograms of MATR1K ... MATR8K~\label{fig:matr}}
\end{figure}

%\end{comment}

>>>>

\section{Preliminary Experiments}

@@

In this section we check if the use of different machines affects the distribution of execution (process) times on the same program. 
In other words, we examine {\em inter-machine repeatability}. For this examination, we use a nested for loop program, here termed INC, 
with different task lengths: 16 seconds, 13 seconds, and 17.2 seconds, each termed {\em INC16}, {\em INC13}, and {\em INC17.2}, respectively.

Figures~\ref{fig:dm_1}, ~\ref{fig:dm_2}, and ~\ref{fig:dm_3} display the process time (PT) histograms on INC16, INC13, and INC17.2 run on our machines (from {\tt sodb8} to {\tt sodb12}), respectively. Based on these figures, we conclude that we should stick to one machine  for consistency, and {\tt sodb9} should be chosen in a sense that the data measured there consistently reveals a distinct binormal pattern.

@@

\subsection{One-to-One Comparison between An Insertion Sort \& A Corresponding INC Programs~\label{sec:sort}} 
This section compares program time histograms 
of an insertion sort and a nested-for-loop programs (termed INC) with different input sizes. The insertion sort program sorts the elements of a given array in non-decreasing order.  The program repeatedly runs 300 times for a given input size. The input size for the program varies from 144K to 344K integer elements, which are randomly generated. Note that each sort program over a specific input size is termed SORT{\it x}: for instance, SORT100 indicates the insertion sort program over 100K elements. An INC program's task length is correspondingly determined by the program time of an SORT program. In Figures~\ref{fig:sort1},~\ref{fig:sort2}, and~\ref{fig:sort3} 
we perform a match on an SORT program and its corresponding INC program.

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort144_dist.eps}
		\label{fig:sort144_dist}
	}	
	\subfigure[PT frequency on INC8.8 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{inc8.8_dist.eps}
		\label{fig:inc8.8_dist}
	}
	\subfigure[PT frequency on SORT177 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort177_dist.eps}
		\label{fig:sort177_dist}
	}	
	\subfigure[PT frequency on INC13 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{inc13.0_dist.eps}
		\label{fig:inc13_dist}
	}
	\caption{PT Histograms I~\label{fig:sort1}}
\end{figure}

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT144 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort244_dist.eps}
		\label{fig:sort144_dist}
	}	
	\subfigure[PT frequency on INC25.8 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{inc25.8_dist.eps}
		\label{fig:inc25.8_dist}
	}
	\subfigure[PT frequency on SORT288  on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort177_dist.eps}
		\label{fig:sort177_dist}
	}	
	\subfigure[PT frequency on INC35 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{inc13.0_dist.eps}
		\label{fig:inc35_dist}
	}
	\caption{PT Histogram Comparison II~\label{fig:sort2}}
\end{figure}

\begin{figure}[H]
	\centering
	\subfigure[PT frequency on SORT344  on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort344_dist.eps}
		\label{fig:sort344_dist}
	}	
	\subfigure[PT frequency on INC51.5 on {\tt sodb10}]{
		\includegraphics[scale=0.43]{inc51.5_dist.eps}
		\label{fig:inc51.5_dist}
	}
	\caption{PT Histogram Comparison III~\label{fig:sort3}}
\end{figure}

@@

\section{Histograms of the Execution Times on Real-World Programs~\label{sec:real-world}} 
This section shows histograms for runs on 
different real-world programs with varying input sizes.  
%We run three kinds of programs: {\it nested for-loop}, {\it insertion sort} and {\it matrix multiplication}.  
In addition to the insertion sort program, we use another kind of program---{\it matrix multiplication}.
For these programs, we varied their input sizes by 2{\small $\times$} while trying some intermediate sizes increased by $\sqrt{2}$ and measured execution times of the programs over each input size. We then exhibit different kinds of histograms in the subsequent sections.

\subsection{Insertion Sort on {\tt sodb9}~\label{sec:new_ins_sort}} 
This section shows a series of histograms of an insertion sorting program that 
sorts the elements of a given array in non-decreasing order. 
The program repeatedly runs 300 times for a given input size. 
Note that each sort program over a specific input size is termed SORT{\it x}: 
for instance, SORT100 indicates the insertion sort program over 100K elements. 
(This experiment was performed on {\tt sodb9}.)

\begin{figure}[htp!]
	\centering
	\subfigure[PT frequency on SORT100 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{s9_sort100_dist.eps}
		\label{fig:s9_sort100_dist}
	}
	\subfigure[PT frequency on SORT200 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{s9_sort200_dist.eps}
		\label{fig:s9_sort200_dist}
	}
	\subfigure[PT frequency on SORT400 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{s9_sort400_dist.eps}
		\label{fig:s9_sort400_dist}
	}
	\subfigure[PT frequency on SORT800 on {\tt sodb9}]{
		\includegraphics[scale=0.43]{s9_sort800_dist.eps}
		\label{fig:s9_sort800_dist}
	}
	\caption{PT Histograms of SORT100 ... SORT800 on {\tt sodb9}~\label{fig:new_s9_sort1}}
\end{figure}

\newpage
\pagebreak

\section{Appendix}

\subsection{Insertion Sort on {\tt sodb8}~\label{sec:add_new}} 
This section shows a series of histograms of an insertion sorting program that 
sorts the elements of a given array in non-decreasing order. 
The program repeatedly runs 300 times for a given input size. 
The input size for the program 
varies from 100,000 to 1,160,000 integer elements, which are randomly generated. 
Note that each sort program over a specific input size is termed SORT{\it x}: 
for instance, SORT100 indicates the insertion sort program over 100K elements. 
This experiment was performed on {\tt sodb8}.

\begin{figure}[h]
	\centering
%	\subfigure[PT frequency on SORT25]{
%		\includegraphics[scale=0.43]{sort25_dist.eps}
%		\label{fig:sort25_dist}
%	}
%	\subfigure[PT frequency on SORT50]{
%		\includegraphics[scale=0.43]{sort50_dist.eps}
%		\label{fig:sort50_dist}
%	}
	\subfigure[PT frequency on SORT100 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort100_dist.eps}
		\label{fig:sort100_dist}
	}
	\subfigure[PT frequency on SORT400 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort400_dist.eps}
		\label{fig:sort400_dist}
	}
	\subfigure[PT frequency on SORT580 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort580_dist.eps}
		\label{fig:sort588_dist}
	}
	\caption{PT Histograms of SORT100 ... SORT580~\label{fig:n_sort1}}
\end{figure}

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on SORT800 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort800_dist.eps}
		\label{fig:sort800_dist}
	}
	\subfigure[PT frequency on SORT1000 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort1000_dist.eps}
		\label{fig:sort1000_dist}
	}
	\subfigure[PT frequency on SORT1160 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort1160_dist.eps}
		\label{fig:sort1160_dist}
	}
	\subfigure[PT frequency on SORT2320 on {\tt sodb8}]{
		\includegraphics[scale=0.43]{sort2320_dist.eps}
		\label{fig:sort2320_dist}
	}
	\caption{PT Histograms of SORT800 ... SORT2320~\label{fig:n_sort2}}
\end{figure}

@@

Figure~\ref{fig:matc} shows a series of 
histograms of the execution times measured on 
the same matrix multiplication program in column major 
as the input sizes grows. The program was run on {\tt sodb10}.

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on MATC1K on {\tt sodb10}]{
		\includegraphics[scale=0.43]{matc1024_dist.eps}
		\label{fig:matc1k_dist}
	}
	\subfigure[PT frequency on MATC2K on {\tt sodb10}]{
		\includegraphics[scale=0.43]{matc2048_dist.eps}
		\label{fig:matc2k_dist}
	}
	\subfigure[PT frequency on MATC4K on {\tt sodb10}]{
		\includegraphics[scale=0.43]{matc4096_dist.eps}
		\label{fig:matc4k_dist}
	}
	\subfigure[PT frequency on MATC8K on {\tt sodb10}]{
		\includegraphics[scale=0.43]{matc8192_dist.eps}
		\label{fig:matc8k_dist}
	}
	\caption{PT Histograms of MATC1024 ... MATC8192~\label{fig:matc}}
\end{figure}

@@

Figure~\ref{fig:matc} shows a series of 
histograms of the execution times measured on 
the same matrix multiplication program in row major as the input sizes grows. 
The program was run on {\tt sodb12}.

@@

\begin{figure}[h]
	\centering
	\subfigure[PT frequency on MATR1K on {\tt sodb12}]{
		\includegraphics[scale=0.43]{matr1024_dist.eps}
		\label{fig:matr1k_dist}
	}
	\subfigure[PT frequency on MATR2K on {\tt sodb12}]{
		\includegraphics[scale=0.43]{matr2048_dist.eps}
		\label{fig:matr2k_dist}
	}
	\subfigure[PT frequency on MATR4K on {\tt sodb12}]{
		\includegraphics[scale=0.43]{matr4096_dist.eps}
		\label{fig:matr4k_dist}
	}
	\subfigure[PT frequency on MATR8K on {\tt sodb12}]{
		\includegraphics[scale=0.43]{matr8192_dist.eps}
		\label{fig:matr8k_dist}
	}
	\caption{PT Histograms of MATR1K ... MATR8K~\label{fig:matr}}
\end{figure}
